---
title: Regression of bugs in GitHub repositories
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup}
repos <- read.csv2("lang_repos.csv")
```

# Process data

Here we calculate the percentage of bugs for each repo, later we use this in
our regression model to see if there is correlation between choice of programming
language and the amount of bugs reported on it.

```{r bug-percent}
# Add bug issues percent and bug pull request percent as columns on every repo
repos$bug_issues_perc <- repos$bug_issues / repos$issues * 100
repos$bug_pull_requests_perc <- repos$bug_pull_requests / repos$pull_requests * 100

# Remove data with malformed relationship between issues and bug issues
repos <- repos[repos$bug_issues_perc != Inf, ]
repos <- repos[!is.na(repos$bug_issues_perc), ]
repos <- repos[repos$bug_issues_perc != 0, ]

# Remove outliers
remove_outliers_from_data_frame <- function(data_frame, x) {
    Q <- quantile(x, probs = c(.25, .75), na.rm = FALSE)
    iqr <- IQR(x)
    return(subset(data_frame, x > (Q[1] - 1.5 * iqr) & x < (Q[2] + 1.5 * iqr)))
}
repos <- remove_outliers_from_data_frame(repos, repos$bug_issues_perc)
```

# Regression

First we calculate mean and median bugs for every language based on the
previously calculated values of the mean and median bugs per repository.
Then we perform regression on the processed data. We use linear regression with
a formulae setting the dependent variable of the mean of bugs on a repository,
independent variables as the language and the amount of stars (popularity).

```{r average-bugs, eval = FALSE}
library(magrittr)
library(tidyverse)
langs_occur <- table(repos$language) # Occurrences of languages in repos

# Calculate mean of mean of repo bugs in repos
lang_bugs_avg <- repos %>%
    group_by(language) %>%
        summarise(sum_bugs = mean(bug_issues_perc))
lang_bugs_avg <- as.data.frame(lang_bugs_avg)

# Calculate mean of median of repo bugs in repos
lang_bugs_median <- repos %>%
    group_by(language) %>%
      summarise(sum_bugs = median(bug_issues_perc))
lang_bugs_median <- as.data.frame(lang_bugs_median)

# Construct data frame for viewing difference in averages
lang_bugs <- data.frame(lang = lang_bugs_avg$language, mean = lang_bugs_avg$sum_bugs, median = lang_bugs_median$sum_bugs)
lang_bugs$occur <- langs_occur
lang_bugs <- lang_bugs[order(lang_bugs$mean, lang_bugs$median, decreasing = TRUE), ]

# Aggregated mean and median bugs
knitr::kable(lang_bugs, col.names = c("Language", "Mean Bug %", "Median Bug %", "Occurrences"))

options("scipen" = 100, "digits" = 4)

# Linear regression of bugs per repo
fit <- lm(repos$bug_issues_perc ~ repos$language + repos$stars)
summary(fit)
```

# Visualisation

```{r visualising}
# Get the number of unique languages in the dataset
n_languages <- length(langs_occur)

# Generate a darker rainbow color palette for the dots
dot_colors <- hcl(seq(15, 375, length.out = n_languages + 1), 100, 30)

library(ggplot2)
library(ggbeeswarm)
library(ggforce)

# Bar chart
ggplot(repos, aes(x = language, y = bug_issues_perc, fill = language)) +
    geom_col() +
    labs(title = "Average Bug Percentage per Language",
         x = "Language",
         y = "Bug Percentage") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_discrete(name = "Language")

# Violin chart
ggplot(repos) +
    aes(x = language, y = bug_issues_perc) +
    geom_violin(fill = "white") +
    geom_sina(aes(color = language)) +
    scale_color_manual(values = dot_colors) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
