---
title: Language popularity regression
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include = FALSE}
lang_repos_clean <- read.csv2("lang_repos_clean_100_000_v2.csv")
```

# Language popularity regression
Here we wish to answer our first research question and attempt a basic prediction
of how language popularity might change in the future. We do this by building
a model using our current data and using it to predict future popularity based
on current trends.

## Preparation of data

We first need to prepare data for analysis, we want to find a regression curve
that can accurately predict the trend of the current data. We will later extend
the model to attempt a basic prediction of future popularity of the top 10
programming languages. In this first step though we only wish to prepare our
data for analysis. Because we wish to predict the trend of language popularity
we need to construct a timeline of popularity for each language. The format of
the data will thus be a data frame containing a column for each language and a
"day" column which will have a row for each day in the timespan. Each language
column will contain the popularity for each of these days. The popularity will
be calculated based on how many of the repositories in the dataset were active
at the time.

An example of our dataset:

| Day | C++ | C# |
|---|---|---|
|2007-10-29|146|10|
|2007-10-30|150|12|

```{r preparing}
library(magrittr)

# Format date strings to date data type
lang_repos_clean$createdAt <-  lang_repos_clean$createdAt %>% as.Date(format = "%Y-%m-%d")
lang_repos_clean$pushedAt <- lang_repos_clean$pushedAt %>% as.Date(format = "%Y-%m-%d")

# Find the start and end date
min <- lang_repos_clean[which.min(lang_repos_clean$createdAt), ]$createdAt
max <- lang_repos_clean[which.max(lang_repos_clean$createdAt), ]$createdAt

# Create sequence timeline spanning timespan
range <- seq(min, max, by = "day")

# Extract the languages that occur in the dataset as unique values
languages <- lang_repos_clean$language %>% unlist() %>% table() %>% names()
top_languages <- lang_repos_clean$language %>% unlist() %>% table() %>% sort(decreasing = TRUE)

# Create timeline for processing language popularity
lang_timeline <- data.frame(day = range)

# Add languages as columns in timeline
lang_df <- data.frame(matrix(0, ncol = length(languages)))
colnames(lang_df) <- languages
lang_timeline <- cbind(lang_timeline, lang_df)

# For every repo add 1 to the timespan where it has been active createdAt -> pushedAt
for (i in 1:nrow(lang_repos_clean)) {
    repo <- lang_repos_clean[i, ]
    lang_timeline[lang_timeline$day %in% repo$createdAt:repo$pushedAt, repo$language] <- 
        lang_timeline[lang_timeline$day %in% repo$createdAt:repo$pushedAt, repo$language] + 1
}

# Remove most recent data as it is inaccurate
lang_timeline <- lang_timeline[lang_timeline$day < max - 500, ]
```

## Regression

Next we perform regression on our timeline, our previous data structure is reshaped
into a simpler data frame using the "melt" function. The data now looks like this:

|day|variable|value|
|---|---|---|
|2007-10-29|C++|146|
|2007-10-30|C++|150|
|2007-10-29|C#|10|
|2007-10-30|C#|12|

This is done to prepare the data for our regression method. We are using the
"gam" function to construct a Gaussian polynomial regression model. This is a model
we landed on after analysing results and visualisations of multiple models.
The method formulae contains a smoothing function to further generalize
results("s" function).

```{r regression}
# Turn timeline language columns into variables on "variable" column
library(reshape2)
lang_time_df <- melt(lang_timeline, id.vars = "day")

# Numerical date representation for use in regression model
lang_time_df$day_num <- as.numeric(lang_time_df$day)

# Regression (GAM) Gaussian polynomial regression model
library(mgcv)
timeline.fit <- gam(value ~ variable + s(day_num, bs = "cs"), data = lang_time_df)

# Remove further unused column
lang_time_df$day_num <- NULL

# Print output in non-scientific notation
options("scipen" = 100, "digits" = 4)

# Print summary of regression model
summary(timeline.fit)

# Select the top 10 programming languages for visualisation
lang_time_df_top <- lang_time_df[lang_time_df$variable %in% names(top_languages[1:10]), ]

# Change name of column "variable"
lang_time_df_top$Language <- lang_time_df_top$variable
lang_time_df_top$variable <- NULL

# Visualising current model and data with the model prediction up to 2030
# Model is generated again internally in graph,
# but uses the same model and formulae as "timeline.fit"
library(ggplot2)
ggplot(lang_time_df_top, aes(x = day, y = value, col = Language)) +
    geom_line() +
    xlim(min, as.Date("2030-01-01")) +
    ylim(0, 14000) +
    xlab("Date") +
    ylab("Popularity") +
    geom_vline(xintercept = max - 500) +
    stat_smooth(
        aes(x = day, y = value, col = Language, linetype = Language),
        method = "gam", formula = y ~ s(x, bs = "cs"),
        fullrange = TRUE) +
    labs(title = "Future trends of the top 10 programming Languages") +
    theme_bw()
```

## Analysis of data
